Police Shootings and Racism in America

### 2020-10-18 police_eda.Rmd
adding eda

all <- all %>% 
  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "tr", "AvgAge", "VA", "F", "nshots", "sit")) %>% 
  mutate(b = as.numeric(b > 0),
         w = as.numeric(w > 0),
         l = as.numeric(l > 0),
         a = as.numeric(a > 0),
         so =  as.numeric(str_count(all$SubjectRace, "O") > 0),
         smale = as.numeric(str_count(all$SubjectGender, "M") > 0),
         sfem = as.numeric(str_count(all$SubjectGender, "F") > 0),
         ob = as.numeric(str_count(all$OfficerRace, "B") > 0),
         ow = as.numeric(str_count(all$OfficerRace, "W") > 0),
         ol = as.numeric(str_count(all$OfficerRace, "L") > 0),
         oa = as.numeric(str_count(all$OfficerRace, "A") > 0),
         oo =  as.numeric(str_count(all$OfficerRace, "O") > 0),
         omale = as.numeric(str_count(all$OfficerGender, "M") > 0),
         ofem = as.numeric(str_count(all$OfficerGender, "F") > 0),
         weapon_other = as.numeric(other), 
         replica = as.numeric(replica), 
         knife = as.numeric(knife), 
         gun = as.numeric(weapon == "gun"), 
         unarmed = as.numeric(weapon == "unarmed"),
         unknown_weapon = as.numeric(weapon == "U"),
         fatal = as.numeric(str_count(all$Fatal, "F") > 0)) %>% 
  select(-c(weapon, SubjectGender, OfficerRace, OfficerGender))

all <- rename(all, sb = b, sw = w, sl = l, sa = a)
```

@@

all$OfficerGender[all$OfficerGender == "FEMALE"] <- "F"
all$OfficerGender[all$OfficerGender == "MALE"] <- "M"
all$OfficerGender[all$OfficerGender == "Male"] <- "M"

```

```{r split_data}
single_victims <- all %>% 
  filter(!str_detect(SubjectRace, ";"))

mult_victims <- all %>% 
  filter(str_detect(SubjectRace, ";"))
```

@@

```{r single_victims-race}
tempsb <- single_victims %>% 
  group_by(sb, fatal) %>% 
  summarise(count = n())

ggplot(tempsb, aes(x = sb, y = count, fill = fatal)) +
  geom_bar(stat="identity") + 
  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings")

tempsw <- single_victims %>% 
  group_by(sw, fatal) %>% 
  summarise(count = n())
tempsa <- single_victims %>% 
  group_by(sa, fatal) %>% 
  summarise(count = n())
tempsl <- single_victims %>% 
  group_by(sl, fatal) %>% 
  summarise(count = n())
tempso <- single_victims %>% 
  group_by(so, fatal) %>% 
  summarise(count = n())

```

@@

>>>>

all <- all %>% 
  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "tr", "AvgAge", "VA", "F", "nshots", "sit")) %>% 
  mutate(b = as.numeric(b > 0),
         w = as.numeric(w > 0),
         l = as.numeric(l > 0),
         a = as.numeric(a > 0),
         so =  as.numeric(str_count(all$SubjectRace, "O") > 0),
         smale = as.numeric(str_count(all$SubjectGender, "M") > 0),
         sfem = as.numeric(str_count(all$SubjectGender, "F") > 0),
         ob = as.numeric(str_count(all$OfficerRace, "B") > 0),
         ow = as.numeric(str_count(all$OfficerRace, "W") > 0),
         ol = as.numeric(str_count(all$OfficerRace, "L") > 0),
         oa = as.numeric(str_count(all$OfficerRace, "A") > 0),
         oo =  as.numeric(str_count(all$OfficerRace, "O") > 0),
         omale = as.numeric(str_count(all$OfficerGender, "M") > 0),
         ofem = as.numeric(str_count(all$OfficerGender, "F") > 0),
         other = as.numeric(other), 
         replica = as.numeric(replica), 
         knife = as.numeric(knife), 
         gun = as.numeric(weapon == "gun"), 
         unarmed = as.numeric(weapon == "unarmed"),
         unknown_weapon = as.numeric(weapon == "U"),
         fatal = as.numeric(str_count(all$Fatal, "F") > 0)) %>% 
  select(-c(weapon, SubjectGender, OfficerGender))

all <- rename(all, sb = b, sw = w, sl = l, sa = a, weapon_other = other)
```

@@

@@

```{r rename 0 and 1 values}
final_eda <- final_df %>%
  mutate(
      fatal = case_when(
        fatal == 1 ~ "Fatal",
        fatal == 0 ~ "Non-Fatal"),
      sb = case_when(
        sb == 1 ~ "Black Victim Present",
        sb == 0 ~ "Black Victim Not Present"
      ),
      sw = case_when(
        sw == 1 ~ "White Victim Present",
        sw == 0 ~ "White Victim Not Present"
      ),
      sl = case_when(
        sl == 1 ~ "Latino Victim Present",
        sl == 0 ~ "Latino Victim Not Present"
      ),
      sa = case_when(
        sa == 1 ~ "Asian Victim Present",
        sa == 0 ~ "Asian Victim Not Present"
      ),
      so = case_when(
        so == 1 ~ "Other Victim Present",
        so == 0 ~ "Other Victim Not Present"
      ),
      smale = case_when(
        smale == 1 ~ "Male Victim Present",
        smale == 0 ~ "Male Victim Not Present"
      ),
      sfem = case_when(
        sfem == 1 ~ "Female Victim Present",
        sfem == 0 ~ "Female Victim Not Present"
      ),
      ob = case_when(
        ob == 1 ~ "Black Officer Present",
        ob == 0 ~ "Black Officer Not Present"
      ),
      ow = case_when(
        ow == 1 ~ "White Officer Present",
        ow == 0 ~ "White Officer Not Present"
      ),
      ol = case_when(
        ol == 1 ~ "Latino Officer Present",
        ol == 0 ~ "Latino Officer Not Present"
      ),
      oa = case_when(
        oa == 1 ~ "Asian Officer Present",
        oa == 0 ~ "Asian Officer Not Present"
      ),
      oo = case_when(
        oo == 1 ~ "Other Officer Present",
        oo == 0 ~ "Other Officer Not Present"
      ),
      omale = case_when(
        omale == 1 ~ "Male Officer Present",
        omale == 0 ~ "Male Officer Not Present"
      ),
      weapon_other = case_when(
        weapon_other == 1 ~ "Other Weapon Present",
        weapon_other == 0 ~ "Other Weapon Not Present"
      ), 
      replica = case_when(
        replica == 1 ~ "Replica Present",
        replica == 0 ~ "Replica Not Present"
      ),
      knife = case_when(
        knife == 1 ~ "Knife Present",
        knife == 0 ~ "Knife Not Present"
      ),
      gun = case_when(
        gun == 1 ~ "Gun Present",
        gun == 0 ~ "Gun Not Present"
      ),
      unarmed = case_when(
        unarmed == 1 ~ "Victim Unarmed",
        unarmed == 0 ~ "Victim Armed"
      ),
      unknown_weapon = case_when(
        unknown_weapon == 1 ~ "Weapon Unknown",
        unknown_weapon == 0 ~ "Weapon Known"
      )
  )

saveRDS(final_eda, 'data/final_eda.rds')
```

```{r race shootings vs population demographics}
most_pop <- tail(names(sort(table(final_eda$city))), 6)

for (c in most_pop) {
  temp <- final_eda %>% 
    filter(city == c)

  vic <- temp %>% 
  summarise(total_b = sum(str_count(SubjectRace, "B")), 
            total_w = sum(str_count(SubjectRace, "W")),
            total_l = sum(str_count(SubjectRace, "L")),
            total_a = sum(str_count(SubjectRace, "A")),
            total_o = sum(str_count(SubjectRace, "O")))

  vic <- vic %>%
    t() %>%
    as.data.frame(stringsAsFactors = F) %>%
    rownames_to_column("value") %>%
    `rownames<-`(NULL)

  demo <-

}

vic <- final_eda %>% 
  summarise(total_b = sum(str_count(SubjectRace, "B")), 
            total_w = sum(str_count(SubjectRace, "W")),
            total_l = sum(str_count(SubjectRace, "L")),
            total_a = sum(str_count(SubjectRace, "A")),
            total_o = sum(str_count(SubjectRace, "O")))
vic <- vic %>%
  t() %>%
  as.data.frame(stringsAsFactors = F) %>%
  rownames_to_column("value") %>%
  `rownames<-`(NULL)

pop <- final_eda %>% 
  summarise(total_b = sum())
```

```{r split into single vs mult}
single_victims <- final_eda %>% 
  filter(NumberOfSubjects <= 1)
nrow(single_victims)

mult_victims <- final_eda %>% 
  filter(NumberOfSubjects > 1)
nrow(mult_victims)
```

```{r single victims race counts and props}
tempsb <- single_victims %>% 
  group_by(sb, fatal) %>% 
  summarise(count = n())
tempsb <- tempsb[3:4,]
tempsb <- rename(tempsb, type = sb)

tempsw <- single_victims %>% 
  group_by(sw, fatal) %>% 
  summarise(count = n())
tempsw <- tempsw[3:4,]
tempsw <- rename(tempsw, type = sw)

tempsa <- single_victims %>% 
  group_by(sa, fatal) %>% 
  summarise(count = n())
tempsa <- tempsa[3:4,]
tempsa <- rename(tempsa, type = sa)

tempsl <- single_victims %>% 
  group_by(sl, fatal) %>% 
  summarise(count = n())
tempsl <- tempsl[3:4,]
tempsl <- rename(tempsl, type = sl)

tempso <- single_victims %>% 
  group_by(so, fatal) %>% 
  summarise(count = n())
tempso <- tempso[3:4,]
tempso <- rename(tempso, type = so)

temp <- rbind(tempsb, tempsw, tempsl, tempsa, tempso) %>% 
  group_by(type) %>% 
  mutate(total = sum(count)) %>% 
  ungroup() %>% 
  mutate(prop = count / total)

race_count <- ggplot(temp, aes(x = type, y = count, fill = fatal)) +
  geom_bar(position = "dodge", stat="identity") + 
  labs(x = "Present Victim Race", y = "Count", title = "Counts for Fatal vs. Non-Fatal Single-Victim Police Shootings Across Different Races") +
  theme(legend.text = element_text(size = 8),
        legend.key.size = unit(0.4, "cm"),
        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
        plot.title = element_text(size=10, hjust = 0.5)) + 
  scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))

race_prop <- ggplot(temp, aes(x = type, y = prop, fill = fatal)) +
  geom_bar(position = "stack", stat="identity") + 
  labs(x = "Present Victim Race", y = "Proportion", title = "Proportions for Fatal vs. Non-Fatal Single-Victim Police Shootings Across Different Races") +
  theme(legend.text = element_text(size = 7),
        legend.key.size = unit(0.4, "cm"),
        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
        plot.title = element_text(size=10, hjust = 0.5)) + 
  scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))

ggsave(filename = "figures/eda1.png", race_count)
ggsave(filename = "figures/eda2.png", race_prop)
```

```{r multiple victims race counts and props}
tempsb <- mult_victims %>% 
  group_by(sb, fatal) %>% 
  summarise(count = n())
tempsb <- tempsb[3:4,]
tempsb <- rename(tempsb, type = sb)

tempsw <- mult_victims %>% 
  group_by(sw, fatal) %>% 
  summarise(count = n())
tempsw <- tempsw[3:4,]
tempsw <- rename(tempsw, type = sw)

tempsa <- mult_victims %>% 
  group_by(sa, fatal) %>% 
  summarise(count = n())
tempsa <- tempsa[3,]
tempsa[2,] <- list("Asian Victim Present", "Fatal", 0)
tempsa <- rename(tempsa, type = sa)

tempsl <- mult_victims %>% 
  group_by(sl, fatal) %>% 
  summarise(count = n())
tempsl <- tempsl[3:4,]
tempsl <- rename(tempsl, type = sl)

temp <- rbind(tempsb, tempsw, tempsl, tempsa) %>% 
  group_by(type) %>% 
  mutate(total = sum(count)) %>% 
  ungroup() %>% 
  mutate(prop = count / total)

race_count1 <- ggplot(temp, aes(x = type, y = count, fill = fatal)) +
  geom_bar(position = "dodge", stat="identity") + 
  labs(x = "Present Victim Race", y = "Count", title = "Counts for Fatal vs. Non-Fatal Multiple-Victim Police Shootings Across Different Races") +
  theme(legend.text = element_text(size = 8),
        legend.key.size = unit(0.4, "cm"),
        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
        plot.title = element_text(size=10, hjust = 0.5)) + 
  scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))

race_prop1 <- ggplot(temp, aes(x = type, y = prop, fill = fatal)) +
  geom_bar(position = "stack", stat="identity") + 
  labs(x = "Present Victim Race", y = "Proportion", title = "Proportions for Fatal vs. Non-Fatal Multiple-Victim Police Shootings Across Different Races") +
  theme(legend.text = element_text(size = 7),
        legend.key.size = unit(0.4, "cm"),
        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
        plot.title = element_text(size=10, hjust = 0.5)) + 
  scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))

ggsave(filename = "figures/eda3.png", race_count1)
ggsave(filename = "figures/eda4.png", race_prop1)
```

@@

all$OfficerGender[all$OfficerGender == "FEMALE"] <- "F"
all$OfficerGender[all$OfficerGender == "MALE"] <- "M"
all$OfficerGender[all$OfficerGender == "Male"] <- "M"
```

### 2020-10-18 submissionii.Rmd
adding eda

```{r num_rows, eval = F}
final_eda <- final_df %>%
  mutate(
      fatal = case_when(
        fatal == 1 ~ "Fatal",
        fatal == 0 ~ "Non-Fatal"
    )
  )

@@

```{r, eval = F}
tempsb <- single_victims %>% 
  group_by(sb, fatal) %>% 
  summarise(count = n())
tempsw <- single_victims %>% 
  group_by(sw, fatal) %>% 
  summarise(count = n())
tempsa <- single_victims %>% 
  group_by(sa, fatal) %>% 
  summarise(count = n())
tempsl <- single_victims %>% 
  group_by(sl, fatal) %>% 
  summarise(count = n())

## why are there no others?
tempso <- single_victims %>% 
  group_by(so, fatal) %>% 
  summarise(count = n())

ggplot(single_victims, aes(x = fatal, y = count)) +
  geom_bar(stat="identity") + 
  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings") 
  group_by(Fatal, race) %>% 
  summarise(count = n())
```

|       Upon first glance, there appears to be 1,883 incidents in the final dataset in which there was only one victim, and 148 with multiple victims. Among these,

@@

- limitations: time series, also counting N;U... combos as N even though some are unknown
## References 
1) Vice 
2) https://www.pnas.org/content/pnas/116/34/16793.full.pdf
3) https://www.kpbs.org/news/2020/jun/18/sweeping-change-us-views-police-violence-new-poll-/
4) http://govred.com/blog/deescalation-training-state-requirements/

>>>>

|       There appears to be 1,906 incidents in the final dataset in which there was only one victim, and 148 with multiple victims.

```{r num_rows, eval = F}
final_eda <- readRDS('data/final_eda.rds')

@@

```{r, figures-race, fig.show="hold", out.width="50%", fig.cap="Counts (left) and Proportions (right) of Different Races Across Single Victim Shootings"}
par(mar = c(2, 2, .1, .1))
knitr::include_graphics("figures/eda1.png")
knitr::include_graphics("figures/eda2.png")
```

|       Among all shootings with only one victim, it appears that the victim is more often Black than some other race. However, it seems that Black victims and most victims of the other race types are more likely to be shot non-fatally than fatally, whereas White victims seem more likely to be shot fatally. Finally, it is important to note that the numbers of victims that are of race types Asian and Other are relatively small. This means that we should proceed with caution when evaluating the effects of each of these races on whether a shooting is fatal or non-fatal.

@@

- limitations: time series, also counting N;U... combos as N even though some are unknown

## References 
1) Vice 
2) https://www.pnas.org/content/pnas/116/34/16793.full.pdf
3) https://www.kpbs.org/news/2020/jun/18/sweeping-change-us-views-police-violence-new-poll-/
4) http://govred.com/blog/deescalation-training-state-requirements/

## Appendix
